---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# micronutr: Utilities for Calculating Indicators of Vitamin and Mineral Status of Populations

<!-- badges: start -->
[![Project Status: WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

Vitamin and mineral deficiencies continue to be a significant public health problem. This is particularly critical in developing countries where deficiencies to vitamin A, iron, iodine, and other micronutrients lead to adverse health consequences. Cross-sectional surveys are helpful in answering questions related to the magnitude and distribution of deficiencies of selected vitamins and minerals. This package provides tools for calculating and determining select vitamin and mineral deficiencies using R.

## Installation

You can install the development version of `micronutr` from [GitHub](https://github.com/nutriverse/micronutr) with:

```{r install, echo = TRUE, eval = FALSE}
if(!require(remotes)) install.packages("remotes")
remotes::install_github("nutriverse/micronutr")
```

## Usage


