---
title: "Determining micronutrient status"
author: Nicholus Tint Zaw

vignette: >
  %\VignetteIndexEntry{Determining micronutrient status}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nutricheckr)
```


## Determining iron storage status

Iron storage status can be determined using serum ferritin concentration. This package includes a set of functions used in determining iron storage status:

1. Functions that identify inflammation status based on the levels of acute-phase proteins;

2. Functions taht correct serum ferritin concentration values based on the inflammation status; and,

3. Functions that categorise iron storage status based on the corrected serum ferritin concentration for a specific target group (either under 5 years old or 5 years and over).

These functions were designed and developed based on iron deficiency status determination described in:

Namaste, S. M., Rohner, F., Huang, J., Bhushan, N. L., Flores-Ayala, R., Kupka, R., Mei, Z., Rawat, R., Williams, A. M., Raiten, D. J., Northrop-Clewes, C. A., & Suchdev, P. S. (2017). Adjusting ferritin concentrations for inflammation: Biomarkers Reflecting Inflammation and Nutritional Determinants of Anemia (BRINDA) project. *The American journal of clinical nutrition*, 106(Suppl 1), 359S–371S. https://doi.org/10.3945/ajcn.116.141762

### Inflammation status

Inflammation, and its different stages, can be identified based on the availability of acute-phase proteins - one of either c-reactive protein (CRP) or α1-acid-glycoprotein (AGP), or both. The detailed classification of inflammation status based on combination of both CRP and AGP apply the following cut-off values:

| **Inflammation Categories** | **Cut-off Points** |
| :--- | :--- |
| Incubation | CRP > 5 mg/L & AGP <= 1 g/L |
| Early convalescence | CRP > 5 mg/L & AGP > 1 g/L |
| Late convalescence | CRP <= 5 mg/L & AGP > 1 g/L |

If only either CRP or AGP is available, the cut-off values for determining inflammation are shown below:

| **Type of Protein** | **Cut-off Points (inflammation)** |
| :--- | :--- |
| CRP | CRP > 5 mg/L |
| AGP | AGP > 1 g/L |

These conventions are used in the set of functions to determine levels of inflammation. For example, to determine inflammation status of an individual with a CRP value of 6 mg/L but with no AGP information available, we use:

```{r mn1, eval = FALSE, echo = TRUE}
detect_inflammation_crp(6)
```

which gives:

```{r mn1a, eval = TRUE, echo = FALSE}
detect_inflammation_crp(6)
```

 Then, in the second stage, based on the inflammation status result from step
 one, there is a specific commend for ferritin value correction and each
 command will provide the new vector called `ferritin_corrected` with the
 fixed ferritin value.
  `correct_ferritin`: use this command if the first stage of identification of
  inflammation was performed based on both acute-phase proteins (CRP and AGP)
  `correct_ferritin_crp`: If the inflammation status of your data was
  identified by CRP protein alone, use this command for correction of ferritin
  value. This command will correct each observation ferritin value respective
  to their inflammation status.
  `correct_ferritin_agp`: If the identification of inflammation status was
  performed by the AGP protein alone, use this command. This command will
  correct the ferritin value based on the respective observation inflammation
  status.
  `correct_ferritin`: If your data observation inflammation status was
  four-category outputs as both CRP and AGP protein identified those, use
  this one. Based on each observation inflammation status, each observation
  ferritin values were corrected.

  If the inflammation status was identified based on combination of both CRP
  and AGP proteins, the ferritin correction values are performed as multiply
  with the correction values provided in the below table.

    | **inflammation Categories** | **Correction Values** |
    | :--- | :--- |
    | Incubation | 0.77 |
    | Early convalescence | 0.53 |
    | Late convalescence | 0.75 |

  If only either CRP or AGP proteins is applied, the different correction
  value are applied based on type of protein as mentioned in the below table.

    | **Type of Protein** | **Correction Values** |
    | :--- | :--- |
    | CRP | 0.65 |
    | AGP | 0.72 |


 In the final stage, based on the corrected ferritin value and specific target
 age group, there were particular commends to perform iron storage status
 diagnosis for individual observation.
  `detect_iron`: If your data had already corrected ferritin value based on
  either acute-phase protein or both, use this command. Based on the age group
  of the individual observation and its inflammation status, this will
  generate the new vector called `iron_storage` and which contains the binary
  outcome of iron storage status ("deficiency" and "no deficiency"). This
  command works for both under 5 and over 5 years old.
  `detect_iron_quali`: If the inflammation status was identified based on the
  qualitative report (such as self-reporting from the survey), use this
  commend for iron status identification. This will also generate the new
  vector called `iron_storage` and which contains the binary outcome of
  iron storage status ("deficiency" and "no deficiency"). But, this can only
  apply to the under 5 years old children population.

  The cut-off points value for identification of iron storage category is
  mentioned in the below table.

    | **Population** | **No Deficiency** |	**Deficiency** |
    | :--- | :--- | :--- | :--- |
    | < 5 year old (both sex) |	Ferritin < 12 |	Ferritin >= 12 |
    | >= 5 years old (Male) |	Ferritin < 15	|  15 >= Ferritin >= 200 |
    | >= 5 years old (Female)	| Ferritin < 15	| 15 >= Ferritin >= 150 |

  If the inflammation status was identified based on qualitative inforomation,
  the cut-off point is applied to original ferritin value as < 30 micrograms
  per liter (µg/l) unit is consider as `deficiency` status.

